# Scenario 1: Change in Land Use {#sec-landuse}

{{< include _setup.qmd >}}

In our first scenario, we changes the land use at The Point.

The change is based on WFRC 2050 SE data forecast, in turn based on development plan of The Point.

This scenario uses the planned residential and commercial development at The Point Utah, the site of the
recently demolished Utah State Prison, to model a major increase in land use.

Near Draper, Utah there is a defunct prison site that is slated for redevelopment (see @fig-draper-prison-location).
The 600-acre site will be developed into a multi-use residential and commercial area, with an emphasis on walkability and transit.
A \acr{BRT} line is planned to run through the development to the Draper FrontRunner station, and across I-15 to Lehi, Utah.
This scenario is to model this development and its effect on transportation demand.

To model this scenario in the trip-based model, the only thing that needed to be done was update the socioeconomic data to reflect the land use changes.
Using the information made publicly available by Point of the Mountain State Land Authority [@state_land_authority_framework_2021], the increase in population and employment was estimated by \acr{WFRC} as shown in @tbl-point-se-data.
The development spans five \acr{TAZ}s (see @fig-prison-new-transit), so the socioeconomic data are spread out across them consistent with the development plan [@cite].
To add this information to the model, the input file containing socioeconomic data for each zone was adjusted so the five zones in the new development had SE data that matched the 2050 estimates for that area.
As seen in @tbl-poin-se-data, approximately 7,430 new households and 22,200 new employments were added to these zones. 

This adds X jobs and X households, etc. (show map/table)

@tbl-the-point-overview shows the change in lu data, and @fig-the-point-zones shows these TAZs.

```{r}
#| label: fig-the-point-zones
#| fig-cap: the point zones

targets::tar_read(lu_plot_new_tazs)
```

```{r}
#| label: tbl-the-point-overview
#| tbl-cap: the point overview

targets::tar_read(lu_new_se_table) %>% 
	kbl()
```

## Scenario Creation

In the WFRC model the change is trivial to implement.
We just copy/pasted the se data from 2050 to 2019 for the affected zones.

This zonal data is used directly in the \acr{WFRC} model, but a new synthetic population is needed for ActivitySim.
We used the zonal data as new targets....

Only created a new synthetic population for the 5 zones

Used census targets from Gateway area in SLC

ActivitySim requires two things for this scenario: updated SE data and a new pop.
The updated SE data is the same as in the WFRC scenario, so is also trivial to implement.
The pop, however, is not.
New controls.

New population for only the affected area, then joined to existing.
There were no existing households so we didn't need to remove any (we're not double-counting).

? Note in realistic scenario you could use just the land use forecast as controls directly, rahter than the whole census shenanigains but we didn't do that.


This is trivial to implement in the WFRC/MAG model, where all that is needed is a change in the TAZ-
level land use data file for the affected TAZs. The new land use data is the WFRC/MAG predicted 2050
data for these zones, as this is when development of the site is scheduled for completion. That is, the 2050
socioeconomic data for these five zones (only) was brought into the 2019 scenario. This added 7,430 new
households and 22,200 new jobs to the area.

In ActivitySim, a modified synthetic population with these new households is needed. Currently, the
synthetic population model (PopulationSim) is set up to use Census tract data for many control variables
including income. At the suggestion of Bert Granberg, the BYU team used the Census tract covering the
Gateway area in SLC to represent the new development patterns at The Point. The income distribution,
etc. of this area will therefore match that of the Gateway area, though the number of households and zonal
SE data will match the WFRC projections for 2050. The new synthetic population was run for this new
development only and added to the base scenario population

## Scenario Analysis

We looked at person-miles for both scenarios.
@fig-lu-personmiles-abm shows the increase in person-miles for the ABM.
Shaded based on "in-zone".
(Not many transit trips since we didn't add transit).
@fig-lu-personmiles-tbm shows for TBM.

In a trip-based model, it is possible to calculate VMT/person-miles traveled/etc. from trips produced in
and/or attracted to the new development zones (see Figure 4). Additionally, Figure 5 shows desire lines
for trips produced in these zones. However, those living in the new development make many more trips
than only those produced in their home zone. Figure 6 shows a desire line plot of all trips taken by the
newly-added households in ActivitySim, and Figure 7 shows the distribution of these trips based on their production zone. Though trip-based models do include non–home-based trips, it is difficult to model
accurately where and by whom these trips are made. Analyzing these trips and capturing the full impact
of new development in this way is only possible in an ABM.
Note that in a trip-based model, a round-trip from home to work and back is regarded as two trips
produced in the home zone. ActivitySim does not deal with productions and attractions in the same way,
so we are counting a trip with an origin or destination in the home zone as produced by that zone e.g. in
Figure 6Figures 6 Figure 7and 7.

```{r}
#| label: fig-lu-personmiles-abm
#| fig-cap: personmiles abm

targets::tar_read(cube_lu_new_pmt_plot)
```

```{r}
#| label: fig-lu-personmiles-tbm
#| fig-cap: personmiles tbm

targets::tar_read(asim_lu_new_pmt_plot)
```

Can't really shade since nhb relocation happens before mode choice.
Best you can do is difference the NHB matrices (@fig-lu-desire-tbm (2-panel)) (If there are nhb trips in/out of zones): clearly some of these are not nhb from residents since they occur in these zones.

An output of the trip-based model is a large trip matrix that contains the number of trips that were produced and attracted between each zone.
Along with this matrix, there are matrices showing productions and attractions based on each mode and each trip purpose.
By selecting just the zones we wanted to analyze, we could see how many trips the new development produced and attracted, and we could see what modes and purposes were connected to those trips.
To visualize this, we used the matrix file and the TAZ shape file to create desire lines as seen in *(some figure we can add)*.

Implications: To compare this to the base scenario, we simply subtracted the base scenario trip matrix by the trip matrix from the land use scenario to get desire lines showing the increase in trips.
As could be reasonably expected, we saw that more trips were produced and attracted by the zones containing the new land use area.
We were also able to get an idea of where each of the trips to and from the area were coming from or going to, and what types of modes where being used.

Talk about the numbers we found and describe the results.

Limitations: While we could see the different trips leaving and entering the new development, we weren't able to see what types of people were making the trips.
We also couldn't tell if those trips were part of a series of trips made by the same person.

```{r}
#| label: fig-lu-desire-tbm
#| fig-cap: lu desire tbm
#| fig-subcap: 
#|   - hb
#|   - nhb

targets::tar_read(cube_lu_new_productions_desire_map)
targets::tar_read(cube_lu_nhb_diff_desire_map)
```

With abm, you can follow people and trace their exact paths (@fig-lu-desire-abm).
Implications: Along with seeing where each trip produced in the new development was going, the activity-based model allowed us to see how trips from the development were related to other trips in a travel tour.
With the synthetic population, we were also able to see which types of people were taking trips from the new development.

```{r}
#| label: fig-lu-desire-abm
#| fig-cap: desire abm

targets::tar_read(asim_lu_new_desire_map)
```
