# Scenario 1: Change in Land Use {#sec-landuse}

{{< include _setup.qmd >}}

The basis for the Land Use scenario is the redevelopment of a derelict prison site near Draper, Utah.
This redevelopment is part of an actual plan for the area, and the new development is known as The Point [@the_point_plan].
The location has high transportation accessibility via I-15 to both Salt Lake City and to Provo and Utah County, and plans to connect the development with fixed-route rapid transit are underway.
The data for the land use changes comes from the WFRC land use forecast, which is in turn based on projections from the Point of the Mountain State Land Authority [@the_point_plan].

## Scenario Creation

Creating the scenario in the WFRC/MAG model is trivial.
No changes other than to the land use/socioeconomic data in these zones were made relative to the baseline scenario.
The Point development is expected to be fully completed by 2050, and its projected land use and socioeconomic data is included in the 2050 WFRC forecast, so the 2050 WFRC land use and socioeconomic data projections are used for this site.
The model uses the land use/socioeconomic data directly, so the only adjustment needed is adding the approximately 7,000 households and 25,000 jobs.

ActivitySim requires two changes for this scenario.
The first change is an update to the TAZ-level land use and socioeconomic data, which is identical to the process for the WFRC model.
The second change is an updated synthetic population to represent the new residents of the affected zones.
In order to keep consistency between model scenarios, a new population was created only for the 5 affected zones and joined to the existing synthetic population.
Creating the new synthetic population followed a similar process as in the baseline scenario (described in the Methods section above), but used the new land use data as new TAZ-level controls.
Many of the controls for PopulationSim use tract-level data from the Census, but existing Census data for The Point site is unrepresentative of the new development, as currently the site lacks residential and economic activity.
Because of this, a Census tract covering part of downtown Salt Lake City is used as a representative substitute.
In other words, the total new households in the area will match the socioeconomic projections of the WFRC/MAG model, but the individuals and households will look like the selected portion of downtown Salt Lake City.
There were no individuals or households in the affected zones in the existing synthetic population, so no individuals or households needed to be removed before joining the two populations.

## Scenario Analysis

Both model types allow for very easy analysis of trip numbers and lengths, as the WFRC model outputs origin-destination trip tables directly by mode and purpose, and ActivitySim outputs a list of trips containing information on origin, destination, and mode.
Figures [-@fig-lu-personmiles-cube] and [-@fig-lu-personmiles-asim], for example, show the new trip-miles produced in the updated zones for the WFRC model and ActivitySim, respectively.
However, there is a crucial difference between the model types, and that is the treatment of trips that do not begin or end at the home.

```{r}
#| label: fig-lu-personmiles-cube
#| fig-cap: Trip-miles produced in the updated zones in the Land Use scenario (WFRC model).
#| fig-env: sidewaysfigure
#| fig-width: 8
#| fig-height: 5

targets::tar_read(cube_lu_new_pmt_plot) +
	scale_fill_bright()
```

```{r}
#| label: fig-lu-personmiles-asim
#| fig-cap: Trip-miles of individuals living in the updated zones (ActivitySim). Many of these trips do not have an origin or destination in the home zone of the individual.
#| fig-scap: Trip-miles of individuals living in the updated zones (ActivitySim).
#| fig-env: sidewaysfigure
#| fig-width: 8
#| fig-height: 5

targets::tar_read(asim_lu_new_pmt_plot) +
	scale_fill_bright()
```

In the WFRC model (and in most trip-based models), homes produce trips with different trip purposes, including Home-based Work, Home-based Other, and Non--home-based trips.
Non--home-based trips can be a significant portion of total travel, as @fig-lu-personmiles-cube shows, but are not as straightforward to model as home-based trips.
Because Non--home-based trips by definition have neither an origin or destination at the home these trips happen exclusively between zones that did not produce them.
There are several methods to redistribute Non--home-based trips in trip-based models.
In the trip generation step of the WFRC/MAG model, households produce Non--home-based trips similarly to any other trip purpose.
However, the trips produced in this step determine only the *quantity* of Non--home-based trips, not the trip ends.
The *distribution* of Non--home-based trips is determined by a trip attraction model.
This distribution is then globally scaled to match the total quantity of Non--home-based trips produced in the trip generation step.
In the end it is believed that the right number of trips are occurring, but these trips are entirely separated from the zones that produced them.

By contrast, ActivitySim models individuals and their travel explicitly, and this makes the treatment of Non--home-based trips much more straightforward.
Each trip is tied to a specific individual with a defined home location, and so no extra "redistribution" step is needed to analyze Non--home-based trips.
In fact, as @fig-lu-personmiles-asim shows, Non--home-based trips can occur as part of any tour type/purpose; there is no separate "Non--home-based" purpose in ActivitySim.
Note that @fig-lu-personmiles-asim counts person-miles by *tour* purpose, using the purposes as defined in ActivitySim, rather than converting the ActivitySim trips to trip purposes in common with the WFRC/MAG model.
ActivitySim makes calculating the VMT associated with the new development --- including trips made by the new residents that are not to or from the new development --- trivial.
With the trip-based model this analysis is much more difficult.
