# Scenario 1: Change in Land Use {#sec-landuse}

{{< include _setup.qmd >}}

The basis for the Land Use scenario is the redevelopment of a derelict prison site near Draper, Utah.
This redevelopment is part of an actual plan for the area, and the new development is known as The Point [@the_point_plan].
The location has high transportation accessibility via I-15 to both Salt Lake City and to Provo and Utah County, and plans to connect the development with fixed-route rapid transit are underway.
The data for the land use changes comes from the WFRC land use forecast, which is in turn based on projections from the Point of the Mountain State Land Authority [@the_point_plan].

The site consists of five zones; @tbl-lu-data shows the households, population, and employment by type of these zones in both scenarios.
Notably, there were no households and relatively few jobs in these zones in the baseline scenario.
No changes other than to the land use/socioeconomic data in these 5 zones were made relative to the baseline scenario.

```{r}
#| label: tbl-lu-data
#| tbl-cap: TAZ-level Socioeconomic Data for Land Use Scenario

by <- targets::tar_read(lu_se_by_table) |> mutate(TAZ = 1:5)
lu <- targets::tar_read(lu_se_lu_table) |> mutate(TAZ = 1:5)

bind_rows(by, lu) %>% 
	skbl(col.names = c("TAZ", "Households", "Population", "Retail", "Industrial", "Other", "Total"),
			 digits = 0) %>% 
	kable_styling() %>% 
	add_header_above(c(" " = 3, "Employment" = 4)) |> 
	pack_rows("Base", 1, 5) |> 
	pack_rows("With the Point", 6, 10) 
```

## Scenario Creation

Creating the scenario in the WFRC / MAG model is trivial.
The Point development is expected to be fully completed by 2050, and its projected land use and socioeconomic data is included in the 2050 WFRC forecast, so the 2050 WFRC land use and socioeconomic data projections are used for this site.
The model uses the land use/socioeconomic data directly, so the only adjustment needed is replacing the data for the specific TAZs with the 2050 forecasted data.
All other TAZs have the same land use data as in the 2019 calibration scenario; effectively, we add the 2050 land use for the relevant zones into the 2019 baseline.

ActivitySim requires two changes for this scenario.
The first change is an update to the TAZ-level land use and socioeconomic data, which is identical to the process for the WFRC model.
The second change is an updated synthetic population to represent the new residents of the affected zones.
In order to keep consistency between model scenarios, a new population was created only for the 5 affected zones and joined to the existing synthetic population.
Creating the new synthetic population followed a similar process as in the baseline scenario (described in the Methods section above), but used the new land use data as new TAZ-level controls.
Many of the controls for PopulationSim use tract-level data from the Census, but existing Census data for The Point site is unrepresentative of the new development, as currently the site lacks residential and economic activity.
Because of this, a Census tract covering part of downtown Salt Lake City is used as a representative substitute.
In other words, the total new households in the area will match the socioeconomic projections of the WFRC / MAG model, but the individuals and households will look like the selected portion of downtown Salt Lake City.
There were no individuals or households in the affected zones in the existing synthetic population, so no individuals or households needed to be removed before joining the two populations.

## Scenario Analysis

There are several kinds of analyses an agency likely would want to do in assessing the effects of a change in land use.
Chief among them would be an analysis of the new trips resulting from the development.
These analyses could include the number of trips, the distance traveled, and where the trips are being made.

Both model types allow for very easy analysis of trip numbers and lengths, as the WFRC model outputs origin-destination trip tables directly by mode and purpose, and ActivitySim outputs a list of trips containing information on origin, destination, and mode.
Figures [-@fig-lu-personmiles-cube] and [-@fig-lu-personmiles-asim], for example, show the new trip-miles produced in the updated zones for the WFRC model and ActivitySim, respectively.
However, there is a crucial difference between the model types, and that is the treatment of trips that do not begin or end at the home.

```{r}
#| label: fig-lu-personmiles-cube
#| fig-cap: Trip-miles produced in the updated zones in the Land Use scenario (WFRC model).
#| fig-env: sidewaysfigure
#| fig-pos: p
#| fig-width: 8
#| fig-height: 5

targets::tar_read(cube_lu_new_pmt_plot) +
	scale_fill_bright()
```

```{r}
#| label: fig-lu-personmiles-asim
#| fig-cap: Trip-miles of individuals living in the updated zones (ActivitySim). Many of these trips do not have an origin or destination in the home zone of the individual.
#| fig-scap: Trip-miles of individuals living in the updated zones (ActivitySim).
#| fig-env: sidewaysfigure
#| fig-pos: p
#| fig-width: 8
#| fig-height: 5

targets::tar_read(asim_lu_new_pmt_plot) +
	scale_fill_bright()
```

In the WFRC model (and in most trip-based models), homes produce trips with different trip purposes, including Home-based Work, Home-based Other, and Non--home-based trips.
"Home-based" trips have an origin or destination at the home, and are fairly straightforward to model, as the destination choice step can take for granted that these trips have one trip end in the zone that produced them.
In addition to home-based trips, though, individuals make many "non--home-based" trips, which do not have an origin or destination at the home (e.g., traveling from work to a grocery store).
Non--home-based trips can be a significant portion of total travel, as @fig-lu-personmiles-cube shows, but are not as straightforward to model as home-based trips.
Because Non--home-based trips by definition have neither an origin or destination at the home (where trips are produced in the trip generation step), these trips happen exclusively between zones that did not produce them.
It is difficult therefore to know how best to redistribute Non--home-based trips, as they could in reality have any number of origins and/or destinations.
Though modeling the destinations for Non--home-based trips could be done via a similar process to that of home-based trips, the origins of these trips need to be modeled as well.
There are several methods to redistribute Non--home-based trips in trip-based models.
Here there are two sources of information for Non--home-based trip ends: a production model and an attraction model.
In the trip generation step of the WFRC / MAG model, households produce Non--home-based trips similarly to any other trip purpose.
However, the trips produced in this step determine only the *quantity* of Non--home-based trips, not the trip ends.
The *distribution* of Non--home-based trips is determined by a trip attraction model, largely based on TAZ employment.
This distribution is then globally scaled to match the total quantity of Non--home-based trips produced in the trip generation step.
In the end it is believed that the right number of trips are occurring, but these trips are entirely separated from the zones that produced them.
No information about their segmentation is therefore available.

By contrast, ActivitySim models individuals and their travel explicitly, and this makes the treatment of Non--home-based trips much more straightforward.
Each trip is tied to a specific individual with a defined home location, and so no extra "redistribution" step is needed to analyze Non--home-based trips: these are "built-in" to each individual's tour pattern.
In fact, as @fig-lu-personmiles-asim shows, Non--home-based trips can occur as part of any tour type/purpose; there is no separate "Non--home-based" purpose in ActivitySim.
Note that @fig-lu-personmiles-asim counts person-miles by *tour* purpose, using the purposes as defined in ActivitySim, rather than converting the ActivitySim trips to trip purposes in common with the WFRC / MAG model.
ActivitySim makes calculating the VMT associated with the new development --- including trips made by the new residents that are not to or from the new development --- trivial.
With the trip-based model this analysis is much more difficult.

This point is perhaps made more clear in the desire line plots in @fig-lu-desire.
@fig-lu-desire-1 shows the home-based trips produced in the zones with the new development, while the the between the Baseline and Point scenarios in the WFRC / MAG model, as the difference between the various home-based trip matrices.
@fig-lu-desire-2, by contrast, shows the complete tours of all the synthetic persons located in the zone.
Missing from these figures are trips made to the zone by non-residents, but in the case of ActivitySim these trips are easily obtained.

```{r, warning=FALSE, message = FALSE}
#| label: fig-lu-desire
#| fig-cap: "Desire lines of trips made by new residents by mode."
#| fig-subcap: 
#|   - "WFRC / MAG Model: Home-based purposes"
#|   - "ActivitySim: All trips"
#| fig-width: 6
#| fig-height: 3
#| layout-nrow: 2
targets::tar_read(cube_lu_new_productions_desire_map)
targets::tar_read(asim_lu_new_desire_map) 
```
