# Scenario 1: Change in Land Use {#sec-landuse}

{{< include _setup.qmd >}}

Near Draper, Utah there is a defunct prison site that is slated for redevelopment (see @fig-draper-prison-location).
The 600-acre site will be developed into a multi-use residential and commercial area, with an emphasis on walkability and transit.
This development is known as The Point, and is the basis for the first model scenario in this paper.

```{r}
#| label: fig-draper-prison-location
#| fig-cap: draper prison location

plot(mtcars[1:3])
```

This scenario models the change in land use that a development such as The Point would create.
Though the actual development plans for The Point include an expansion of transit services, only the additional households and jobs created from this development are modeled in this scenario.
The data for the land use changes comes from the WFRC land use forecast, which is in turn based on projections from the Point of the Mountain State Land Authority [@state_land_authority_framework_2021].
The development is expected to be fully completed by 2050, and so the 2050 WFRC land use and socioeconomic data projections are used for this site.

The site consists of 5 TAZs, as shown in @fig-the-point-zones.
@tbl-the-point-data shows the additional households, population, and employment by type added to each TAZ in this scenario.
***Notably, there were no households or jobs in these TAZs in the baseline scenario, so @tbl-the-point-data also represents the total in each category.***\
No changes other than to the land use/socioeconomic data in these 5 TAZs was made relative to the baseline scenario.

```{r}
#| label: fig-the-point-zones
#| fig-cap: the point zones

targets::tar_read(lu_plot_new_tazs)
```

```{r}
#| label: tbl-the-point-data
#| tbl-cap: the point data

targets::tar_read(lu_new_se_table) %>% 
	kbl()
```

## Scenario Creation

In the WFRC model, this change is trivial to implement.
The model uses the land use/socioeconomic data directly, so the only adjustment needed is replacing the data for the specific TAZs with the 2050 data.
As noted previously, all other TAZs have the same land use data as in the 2019 baseline scenario.

ActivitySim requires two changes for this scenario.
The first is an update to the TAZ-level land use and socioeconomic data, which is identical to the process for the WFRC model.
The second is an updated synthetic population.
In order to keep consistency between model scenarios, a new population was created only for the 5 affected TAZs and joined to the existing synthetic population.
There were no individuals or households in the affected zones in the existing synthetic population, so no individuals or households needed to be removed before joining the two populations.

Creating the new synthetic population followed a similar process as in the baseline scenario (@sec-populationsim), but used the new land use data as new TAZ-level controls.
However, many of the controls for PopulationSim use tract-level data from the Census, and existing Census data for these controls is unrepresentative of the new development.
Because of this, the Census tract covering the Gateway area in Salt Lake City (see @fig-gateway-tract) is used to represent the new development patterns at The Point.
The income distribution, etc. of this area will therefore match that of the Gateway area, though the TAZ-level controls and land use/socioeconomic data will match the WFRC projections for 2050.

? Note in realistic scenario you could use just the land use forecast as controls directly, rahter than the whole census shenanigans but we didn't do that since we wanted the models to be independent.

## Scenario Analysis

We looked at person-miles for both scenarios.
@fig-lu-personmiles-abm shows the increase in person-miles for the ABM.
Shaded based on "in-zone".
(Not many transit trips since we didn't add transit).
@fig-lu-personmiles-tbm shows for TBM.

In a trip-based model, it is possible to calculate VMT/person-miles traveled/etc.
from trips produced in and/or attracted to the new development zones (see Figure 4).
Additionally, Figure 5 shows desire lines for trips produced in these zones.
However, those living in the new development make many more trips than only those produced in their home zone.
Figure 6 shows a desire line plot of all trips taken by the newly-added households in ActivitySim, and Figure 7 shows the distribution of these trips based on their production zone.
Though trip-based models do include non–home-based trips, it is difficult to model accurately where and by whom these trips are made.
Analyzing these trips and capturing the full impact of new development in this way is only possible in an ABM.
Note that in a trip-based model, a round-trip from home to work and back is regarded as two trips produced in the home zone.
ActivitySim does not deal with productions and attractions in the same way, so we are counting a trip with an origin or destination in the home zone as produced by that zone e.g. in Figure 6Figures 6 Figure 7and 7.

```{r}
#| label: fig-lu-personmiles-abm
#| fig-cap: personmiles abm

targets::tar_read(cube_lu_new_pmt_plot)
```

```{r}
#| label: fig-lu-personmiles-tbm
#| fig-cap: personmiles tbm

targets::tar_read(asim_lu_new_pmt_plot)
```

Can't really shade since nhb relocation happens before mode choice.
Best you can do is difference the NHB matrices (@fig-lu-desire-tbm (2-panel)) (If there are nhb trips in/out of zones): clearly some of these are not nhb from residents since they occur in these zones.

An output of the trip-based model is a large trip matrix that contains the number of trips that were produced and attracted between each zone.
Along with this matrix, there are matrices showing productions and attractions based on each mode and each trip purpose.
By selecting just the zones we wanted to analyze, we could see how many trips the new development produced and attracted, and we could see what modes and purposes were connected to those trips.
To visualize this, we used the matrix file and the TAZ shape file to create desire lines as seen in *(some figure we can add)*.

Implications: To compare this to the base scenario, we simply subtracted the base scenario trip matrix by the trip matrix from the land use scenario to get desire lines showing the increase in trips.
As could be reasonably expected, we saw that more trips were produced and attracted by the zones containing the new land use area.
We were also able to get an idea of where each of the trips to and from the area were coming from or going to, and what types of modes where being used.

Talk about the numbers we found and describe the results.

Limitations: While we could see the different trips leaving and entering the new development, we weren't able to see what types of people were making the trips.
We also couldn't tell if those trips were part of a series of trips made by the same person.

```{r}
#| label: fig-lu-desire-tbm
#| fig-cap: lu desire tbm
#| fig-subcap: 
#|   - hb
#|   - nhb

targets::tar_read(cube_lu_new_productions_desire_map)
targets::tar_read(cube_lu_nhb_diff_desire_map)
```

With abm, you can follow people and trace their exact paths (@fig-lu-desire-abm).
Implications: Along with seeing where each trip produced in the new development was going, the activity-based model allowed us to see how trips from the development were related to other trips in a travel tour.
With the synthetic population, we were also able to see which types of people were taking trips from the new development.

```{r}
#| label: fig-lu-desire-abm
#| fig-cap: desire abm

targets::tar_read(asim_lu_new_desire_map)
```
