# Scenario 1: Change in Land Use {#sec-landuse}

{{< include _setup.qmd >}}

Near Draper, Utah there is a defunct prison site that is slated for redevelopment.
The 600-acre site will be developed into a multi-use residential and commercial area, with an emphasis on walkability and transit.
This development is known as The Point, and is the basis for the first model scenario in this paper.
@fig-draper-prison-location shows the location of this development relative to the surrounding area.

```{r}
#| label: fig-draper-prison-location
#| fig-cap: '"The Point" development site (shown in blue).'

targets::tar_read(draper_prison_site_map)
```

This scenario models the change in land use that a development such as The Point would create.
Though the actual development plans for The Point include an expansion of transit services, only the additional households and jobs created from this development are modeled in this scenario.
The data for the land use changes comes from the WFRC land use forecast, which is in turn based on projections from the Point of the Mountain State Land Authority [@state_land_authority_framework_2021].
The development is expected to be fully completed by 2050, and so the 2050 WFRC land use and socioeconomic data projections are used for this site.

The site consists of 5 TAZs, as shown in @fig-the-point-zones.
@tbl-the-point-data shows the households, population, and employment by type of these TAZs in both the baseline scenario and with the new land use.
Notably, there were no households and relatively few jobs in these TAZs in the baseline scenario.
No changes other than to the land use/socioeconomic data in these 5 TAZs was made relative to the baseline scenario.

```{r}
#| label: fig-the-point-zones
#| fig-cap: Map of each TAZ in The Point development.

targets::tar_read(lu_plot_new_tazs)
```

```{r}
#| label: TAZ-level Socioeconomic Data for The Point
#| tbl-cap: the point data
#| layout-ncol: 2
#| tbl-subcap: 
#|   - Baseline Scenario
#|   - New Land Use

targets::tar_read(lu_se_by_table) |>
	flextable()
targets::tar_read(lu_se_lu_table) |> 
	flextable()
```

## Scenario Creation

In the WFRC model, this change is trivial to implement.
The model uses the land use/socioeconomic data directly, so the only adjustment needed is replacing the data for the specific TAZs with the 2050 data.
As noted previously, all other TAZs have the same land use data as in the 2019 baseline scenario.

ActivitySim requires two changes for this scenario.
The first is an update to the TAZ-level land use and socioeconomic data, which is identical to the process for the WFRC model.
The second is an updated synthetic population.

In order to keep consistency between model scenarios, a new population was created only for the 5 affected TAZs and joined to the existing synthetic population.
There were no individuals or households in the affected zones in the existing synthetic population, so no individuals or households needed to be removed before joining the two populations.

Creating the new synthetic population followed a similar process as in the baseline scenario (@sec-populationsim), but used the new land use data as new TAZ-level controls.
However, many of the controls for PopulationSim use tract-level data from the Census, and existing Census data for these controls is unrepresentative of the new development, due to the lack of households and economic activity at the site of the development.
Because of this, the Census tract covering the Gateway area in Salt Lake City (see @fig-gateway-tract) is used to represent the new development patterns at The Point.
The income distribution, etc. of this area will therefore match that of the Gateway area, though the TAZ-level controls and land use/socioeconomic data in the area will match the WFRC projections for 2050.

```{r}
#| label: fig-gateway-tract
#| fig-cap: Census tract for the Gateway area in Salt Lake City (shown in blue). Tract-level controls for the synthetic population in the new development come from this tract rather than the tract at the development site.
#| fig-scap: Census tract for the Gateway area in Salt Lake City.

plot(mtcars[1:3])
```

In a more realistic case, a transportation agency would have forecasted land use and socioeconomic data that could be used as controls to PopulationSim, rather than using a (different) representative Census tract for new development.
However, our ActivitySim implementation only needs to be within a rough approximation of the WFRC model for the purposes of this project, and the method used here results in reasonable accuracy between the models.
<!-- @tbl-new-land-use-comparison shows that the new synthetic population has ***only minor discrepancies  --> Additionally, our ActivitySim implementation is designed to be independent from the WFRC model where feasible.

## Scenario Analysis

There are several kinds of analyses an agency likely would want to do in assessing the effects of a change in land use.
Probably chief among them would be an analysis of the new trips resulting from the development.
This could include analysis in the number of trips, the distance traveled, and where the trips are being made.

Both model types allow for very easy analysis of trip numbers and lengths, as the WFRC model outputs origin-destination trip tables directly by mode and purpose, and ActivitySim outputs a list of trips containing information on origin, destination, and mode.
[@fig-lu-personmiles-cube; @fig-lu-personmiles-abm], for example, show the new trip-miles produced in the updated zones for the WFRC mode and ActivitySim, respectively.
However, there is a crucial difference between the model types, and that is the treatment or trips that do not begin or end at the home.

```{r}
#| label: fig-lu-personmiles-cube
#| fig-cap: personmiles cube

targets::tar_read(cube_lu_new_pmt_plot)
```

```{r}
#| label: fig-lu-personmiles-asim
#| fig-cap: personmiles asim

targets::tar_read(asim_lu_new_pmt_plot)
```

In the WFRC model (and in many trip-based models), homes produce trips with different trip purposes, including home-based work, home-based other, and non--home-based trips.
"Home-based" trips have an origin or destination at the home, and are fairly straightforward to model, as the destination choice step can take for granted that these trips have one trip end in the zone that produced them.
In addition to home-based trips, though, individuals make many "non--home-based" trips, which do not have an origin or destination at the home (e.g. traveling from work to a grocery store).
Non--home-based trips can be a significant portion of total travel, as @fig-lu-personmiles-cube shows, but are not as straightforward to model as home-based trips.

Because non--home-based trips by definition have neither an origin or destination at the home (where trips are produced in the trip generation step), these trips happen exclusively between zones that did not produce them.
It is difficult therefore to know how best to redistribute non--home-based trips, as they could in reality have any number of origins and destinations.
Though modeling the destination choice for non--home-based trips could be done via a similar process to that of home-based trips, the origins of these trips need to be modeled as well.

There are several methods to redistribute non--home-based trips in trip-based models.
One naïve approach is to assign non--home-based trip origins in a similar manner to trip destinations as part of the trip distribution step, and then assign destinations to those trips normally *(I don't know if this is actually a thing, I just made it up)*.
***paragraph about this, cite some sources***.

By contrast, an ABM models individuals and their travel explicitly, and this makes the treatment of non--home-based trips much more straightforward.
Each trip is tied to a specific individual with a defined home location, and so no extra "redistribution" step is needed to analyze non--home-based trips: these are "built-in" to each individual's tour pattern.
In fact, as @fig-lu-personmiles-asim shows, non--home-based trips can occur as part of any tour type/purpose; there is no separate "non--home-based" purpose in ActivitySim.
Note that @fig-lu-personmiles-asim counts person-miles by *tour* purpose, using the purposes as defined in ActivitySim, rather than converting the ActivitySim trips to the "common" trip purposes as discussed in @sec-baseline-calibration.

In addition to looking at total person-miles traveled, it is also useful to analyze where the new trips are going.
One common way to visualize this is with desire lines, which show lines for each trip origin/destination pair.
The thickness of the line represents the number of trips between the pair of zones.

@fig-lu-desire-cube-hb shows a desire line plot by mode for ***all?*** home-based trips.

@fig-lu-desire-cube shows desire line plots of trips generated by the new development zones.
For home-based trips, it is easy to see where trips are produced and attracted to, and this is shown in @fig-lu-desire-cube-1.
However, the non--home-based trips are more difficult to analyze.
Because non--home-based trips are relocated to zones other than where they were generated, it is impossible to filter to only those non--home-based trips generated by the new zones.

Notably, the WFRC model does not redistribute non--home-based trips at all, so these are treated the same as home-based trips for destination choice, mode choice, and network assignment.
In other words, non--home-based trips in the WFRC model have an origin or destination at home, and are functionally "home-based other" trips.
This explains why @fig-lu-desires-cube-2 shows that the increase in non--home-based trips from the baseline scenario consists entirely of trips with an origin or destination in the new development.

The best approach is to take the difference in non--home-based trips between the scenarios (@fig-lu-desire-cube-2), though this presents two problems.
The first is that for real-world analyses, where land use changes are not isolated, it becomes impossible to tell which non--home-based trips are generated by which developments, as the trips are seen only in aggregate.
The second problem is with the distribution of the non--home-based trips.
Though the exact distribution process depends on the specific model used, @fig-lu-desire-cube-2 shows inconsistencies with the approach in the WFRC model.
Between the baseline scenario and the updated land use scenario, many pairs of zones saw an increase in the number of non--home-based trips between them, but other pairs saw a *decrease*.
Additionally, all pairs of zones that saw an increase in non--home-based trips include a production or attraction to the new development zones.
***The WFRC model redistributes non--home-based trips as part of its network assignment step, which occurs*** **after** ***all trip matrices are created***.
*it's either this or the only increase in nhb trips is from non-residents, which doesn't make much sense*.
The only measure of the increase of non--home-based trips after redistribution is therefore in the highway network, reported as roadway volumes.
This compounds with the first problem and makes it even more difficult to tell where the non--home-based trips are coming from.

```{r}
#| label: fig-lu-desire-cube
#| fig-cap: "Desire lines of trips made in the WFRC model by mode. Note that the non\u2013home-based trips are obtained by differencing the non\u2013home-based trip matrix with the base year."
#| fig-scap: "Desire lines of trips made in the WFRC model by mode."
#| fig-subcap: 
#|   - "Trips produced in the updated TAZs."
#|   - "Difference in non\u2013home-based trips."

targets::tar_read(cube_lu_new_productions_desire_map)
targets::tar_read(cube_lu_nhb_diff_desire_map)
```

As mentioned, an ABM allows for tracking of individuals explicitly, and so analyzing non--home-based trips is much more straightforward.
@fig-lu-desire-abm shows desire lines of all trips made by individuals living in the new development zones, colored by place of production.
It is also easy to see how trips are related to each other, as each individual has a specific sequence of trips.
The individual nature of an ABM avoids entirely the problems trip-based models have with non--home-based trips.
In a complicated land use forecast, each development's full contribution to network congestion can be analyzed individually.

<!-- With the synthetic population, we were also able to see which types of people were taking trips from the new development. -->

```{r}
#| label: fig-lu-desire-asim
#| fig-cap: "Desire lines of trips made in ActivitySim by mode."

targets::tar_read(asim_lu_new_desire_map)
```
