# Transit {#sec-transit}

{{< include _setup.qmd >}}

The second sceario updates the transit network (frontrunner).

Data comes from the WFRC 2050 build plan.
Includes new stations, new headways, and faster speeds.

Show map of changes.

```{r}
targets::tar_read(frontrunner_plot)
```

## Creating scenario

WFRC model needs three adjustments: edit the headways, point the model to the new frontrunner network (which includes speeds), and add park-and-ride option at new station locations in highway?
network.

Asim just needs new skims, gotten from the new WFRC scenario.

## Analyses

```{r}
targets::tar_read(combined_tr_mode_split_diff) %>% 
	kbl()
```

```{r}
#| fig-cap: needs to be redone?
targets::tar_read(asim_tr_mode_switching_plot)
```

\

```{r}
#| tbl-cap: cube
targets::tar_read(cube_tr_productions_se_summary) %>% 
	kbl()
```

```{r}
#| tbl-cap: asim
targets::tar_read(asim_tr_trips_se_summary) %>% 
	kbl()
```

```{r}
targets::tar_read(combined_tr_new_transit_income_plot)
```

\

```{r}
#| fig-cap: cube
targets::tar_read(cube_tr_diff_by_district_plot)
```

```{r}
#| fig-cap: asim
targets::tar_read(asim_tr_diff_by_district_plot)
```
