# Scenario 3: Increase in Remote Work {#sec-wfh}

{{< include _setup.qmd >}}

With the onset of the COVID 19 pandemic, there were unprecedented numbers of people working remotely [@bick_work_2021].
Though remote work is currently not as common as during the pandemic, remote work rates are increasing each year and are predicted to continue to rise [@cite].
The WFRC model predicts a year-over-year increase in remote work rates of about XXX%, as illustrated in @fig-wfrc-remote-work-rate-plot.

```{r}
#| label: fig-wfrc-remote-work-rate-plot
#| fig-cap: wfrc remote work plot

plot(mtcars[1:3])
```

This scenario is a "what-if" analysis that models a significant increase in remote work rates.
We use the WFRC model predicted remote work rates in the year 2050, but make no other changes from the baseline scenario.
In other words, this scenario models the 2050 predicted remote work rates with the 2019 land use and infrastructure.

## Considerations for Modelling Remote Work ***This section maybe needs to go somewhere else*** (probably lit review)

There has been much research, especially in recent years, on the implications of remote work.
While many MPOs have adjusted their models to account for remote work, and most models follow similar principles, it is not obvious what the best method is. @bramberga_teleworking_2023 even suggested that considerations for remote work should be made on a case by case basis because there is no single best approach.

Increasing remote work rates may affect several aspects of travel behavior.
The most obvious effect is that people will on average make fewer work trips, and this effect will vary by job type [@yasenov_who_2020].
Most travel demand models include a decrease in work trips based on remote work rates and job type [@bramberga_teleworking_2023; @moeckel_working_2017; @sener_copula-based_2011].

While work trips decrease with an increase in remote work, @kim_is_2017 discusses a "rebound effect", where individuals make more discretionary trips on days they do not commute to work.
[@moreno_microscopic_2017] similarly proposes the existence of a "travel time budget", where an increase in trips of one purpose will decrease the time people allocate for trips of another purpose.

This rebound effect is not straightforward, however.
@ellder_telework_2020, for example, finds that distinguishing between people that work from home all day and part of the day might make a difference.
Compared to those who commute to work, the Sweden study shows that those who worked from home the whole day had less trips and miles traveled, but those who worked from home only part of the day had more trips and miles traveled.

Additionally, the types of trips people make can differ depending on remote work status.
While the rebound effect proposes that the *number* of trips may increase on remote work days, @mokhtarian_the_1998 finds a decrease in vehicle *miles* traveled for both work and discretionary trips on remote work days.
This implies that longer trips are being replaced by shorter trips on days people do not travel to work.
@moeckel_working_2017 additionally finds that those who travel to their job site less frequently are more likely to live further away from their job site, and so their longer but infrequent commute is dropped on remote work days, perhaps in favor of shorter, discretionary trips.

<!-- It is also not clear if the rebound effect is real. @zhu_impact_2014 find that there are more miles traveled on the days people telecommute, but @mokhtarian_the_1998 saw opposite effects where vehicle miles traveled decreased for both work and discretionary trips on telecommuting days. -->

<!-- Using data from the 2007 Chicago Regional Household Travel Inventory, @he_telecommuting_2015 found that, after adjusting for other factors like income, household size, and job industry, people who telecommute frequently make 0.072 more discretionary trips on average each day compared to those who don't telecommute. -->

## Scenario Creation

As noted in @sec-baseline-calibration, both models make a distinction between "working from home" (no work location other than home) and "telecommuting" (working remotely some but not all days).
The WFRC model contains a lookup table of both work-from-home (called "home-based jobs" in the WFRC model) and telecommute percentages by job type and year.
Two changes are needed in the WFRC model for this scenario.
The first is to replace the 2019 estimates for work-from-home and telecommuting with the 2050 estimates.
@tbl-wfrc-remote-work-years shows both the original and updated estimates.
The second change is to the TAZ-level socioeconomic data.
The WFRC model estimates a number of home-based jobs in each TAZ, and the original home-based job estimates are similarly replaced with the 2050 estimates.

The WFRC model additionally includes a global scaling factor for all remote work percentages.
However, this was left unchanged (set at 1) as we considered that the 2050 predicted remote work percentages would better model a more realistic increase in remote work than simply adjusting the 2019 rates globally.

```{r}
#| label: tbl-wfrc-remote-work-years
#| tbl-cap: wfrc remote work data

targets::tar_read(cube_wfh_by_remote_work_pct_comparison) %>% 
	kbl()
```

We adjusted the remote work models in ActivitySim using the same process as in @sec-baseline-calibration, but with the 2050 targets from the WFRC model.
The "target work-from-home percent" value in ActivitySim's work-from-home submodel was changed to 3.5% based on a weighted average from the 2050 WFRC data, and the job type coefficients in the telecommute frequency submodel were calibrated to match the WFRC target telecommute shares by job type.
@tbl-wfh-telecommute shows the WFRC 2050 telecommute percentages with the ActivitySim telecommute utility coefficients.
As in the baseline scenario, this calibration allowed ActivitySim to match the WFRC telecommute percentages exactly.

```{r}
#| label: tbl-wfh-telecommute
#| tbl-cap: wfh telecommute

targets::tar_read(comparison_wfh_telecommute_coeffs) %>%
  mutate(pct = scales::label_percent(accuracy = 0.01)(pct)) %>%
  rename(
    "Job Type" = name,
    "WFRC Telecommute %" = pct
    ) %>%
  kbl(
    booktabs = TRUE,
    align = "c"
  ) %>%
  add_header_above(c(" " = 2, "ActivitySim Telecommute Utility Coefficients" = 3)) %>%
  kable_styling()
```

## Scenario Analysis

We can see (@tbl-wfh-mode-split-comp) that asim has more HBO trips and cube does not.
This is because asim lets you take a non-mandatory tour if you work from home, and cube just cuts off hbw trips.

An additional note about the WFH implementation in ActivitySim is that because it is done before tour generation, an individual who does not need to make a work tour may be free to make other discretionary tours, and so this does at least somewhat model an increase in other trip types coinciding with increased telecommuting.

The WFRC model does not include a potential increase in discretionary trips for those who work remotely, though some literature suggests this may happen in reality.
If there was more guidance in the literature on how telecommuting affects other types of trips, then we could have tried to represent that in the model by adding a certain amount of discretionary trips according to the telecommute percentages.
With the \acr{WFRC} model, it might be difficult to have the additional discretionary trips be coming from the same zones as the telecommute trips.
*(I'm not sure about this. I'll have to look more into it)*

```{r}
#| label: tbl-wfh-mode-split-comp
#| tbl-cap: wfh mode split comp

targets::tar_read(combined_wfh_mode_split_diff) %>% 
	kbl()
```

Using the outputs from the trip-based model that show lengths and number of trips between zones, we calculated the vehicle miles traveled for each trip.
This allowed us to see which zones had the biggest decrease in VMT.
This also allowed us to find the number of trips of certain length to create a trip length frequency distribution that can be seen in (*another figure*).

```{r}
#| fig-cap: cube

targets::tar_read(cube_wfh_trip_pmt_diff) %>% 
	kbl()
```

```{r}
#| fig-cap: asim

targets::tar_read(asim_wfh_trip_pmt_diff) %>% 
	kbl()
```

It is straightforward to determine decreased VMT by production zone (and similar analyses) in both models (see Figure 11Figures 11 andFigure 12 12).
Additionally, the team made a desire line plot in the WFRC model of decreased trips.
This, however, is essentially a scaled-down version of the total trips in the Base scenario (see Figure 13).
Remote work percentages do vary by industry, but there is no accounting for trips lengths or personal variables.
Figure 14 shows these same plots for ActivitySim.
At first glance, this also seems like a scaled-down version of the total trips, even though ActivitySim’s remote work models do include distance to work and personal attributes.
It is difficult to tell from the figure alone.
The team plans to do more analysis on this point.

Include a table of the asim tc and wfh coefficients.
Maybe put it in the methodology.

```{r}
#| fig-cap: asim
targets::tar_read(asim_wfh_tlfd_diff_comp_plot)
```

```{r}
#| fig-cap: cube
targets::tar_read(cube_wfh_tlfd_diff_comp_plot)
```

```{r}
targets::tar_read(combined_wfh_tlfd_diff_plot)
```
