# Scenario 1: Land Use

Near Draper, Utah there is a defunct prison site that is slated for redevelopment (see @fig-draper-prison-location).
The 600-acre site will be developed into a multi-use residential and commercial area, with an emphasis on walkability and transit.
A \acr{BRT} line is planned to run through the development to the Draper FrontRunner station, and across I-15 to Lehi, Utah.
This scenario is to model this development and its effect on transportation demand.

To model this scenario in the trip-based model, the only thing that needed to be done was update the socioeconomic data to reflect the land use changes.
Using the information made publicly available by Point of the Mountain State Land Authority [@state_land_authority_framework_2021], the increase in population and employment was estimated as shown in @tbl-point-se-data.
It is important to note that the development spans five \acr{TAZ}s (see @fig-prison-new-transit), and so the socioeconomic data are spread out across them consistent with the development plan [@cite].
To add this information to the model, the csv file containing the socioeconomic data for each zone needed to be adjusted.
This involved simply adding the new information to each of the five rows that represented the zones in the new development. 

```{r}
#| label: fig-draper-prison-location
#| fig-cap: draper prison location

boxplot(mtcars$mpg ~ mtcars$cyl)
```

```{r}
#| label: tbl-point-se-data
#| tbl-cap: the point se data

knitr::kable(head(mtcars))
```

This zonal data is used directly in the \acr{WFRC} model, but a new synthetic population is needed for ActivitySim.
We used the zonal data as new targets....