# Scenario 1: Land Use

Near Draper, Utah there is a defunct prison site that is slated for redevelopment (see @fig-draper-prison-location).
The 600-acre site will be developed into a multi-use residential and commercial area, with an emphasis on walkability and transit.
A \acr{BRT} line is planned to run through the development to the Draper FrontRunner station, and across I-15 to Lehi, Utah.
This scenario is to model this development and its effect on transportation demand.

To model this scenario in the trip-based model, the only thing that needed to be done was update the socioeconomic data to reflect the land use changes.
Using the information made publicly available by Point of the Mountain State Land Authority [@state_land_authority_framework_2021], the increase in population and employment was estimated by \acr{WFRC} as shown in @tbl-point-se-data.
The development spans five \acr{TAZ}s (see @fig-prison-new-transit), so the socioeconomic data are spread out across them consistent with the development plan [@cite].
To add this information to the model, the input file containing socioeconomic data for each zone was adjusted so the five zones in the new development had SE data that matched the 2050 estimates for that area.
As seen in @tbl-poin-se-data, approximately 7,430 new households and 22,200 new employments were added to these zones. 

```{r}
#| label: fig-draper-prison-location
#| fig-cap: draper prison location

boxplot(mtcars$mpg ~ mtcars$cyl)
```

```{r}
#| label: tbl-point-se-data
#| tbl-cap: the point se data

knitr::kable(head(mtcars))
```

This zonal data is used directly in the \acr{WFRC} model, but a new synthetic population is needed for ActivitySim.
We used the zonal data as new targets....

Only created a new synthetic population for the 5 zones

Used census targets from Gateway area in SLC
@fig-point-hbw-productions
# Plots made by Kamryn to put somewhere

```{r}
#| label: fig-point-hbw-productions
#| fig-cap: Additional HBW Productions at "The Point"

targets::tar_read(lu_hbw_trip_diff)

```
