# Methodology {#sec-methods}

```{r}
#| label: setup
#| include: false

library(kableExtra)
```


This section lists our methodology.

<!-- We will compare models... -->

<!-- We will forecast several scenarios. -->

In order to compare trip-based and activity-based models, this research takes a specific model from both types to be representative.
The two models that this research compares are the \acr{WFRC} travel demand model (trip-based) and ActivitySim (activity-based).

## WFRC Model

The \acr{WFRC} model is implemented in the CUBE software by Bentley [@bentley_systems_cube], and is currently used by \acr{WFRC} for modeling travel in the Salt Lake City, Utah area.
\acr{WFRC} provided the model directly, including land use forecasts and the current long-range transportation plan.
The model is taken mostly as-is, though some changes were made to better accommodate the different scenarios.
These modifications are noted in more detail in their respective sections.

The \acr{WFRC} model, like many trip-based models, requires the following inputs:
- **Land use data.** This includes information about population, employment, and socioeconomic variables such as income, delineated by \acr{TAZ}.
In our case this was provided by \acr{WFRC} directly, as an output of their land use forecasting model(s).
- **Travel skims.** This details travel time, cost, etc. between each origin-destination pair of \acr{TAZ}s.
The \acr{WFRC} model uses an iterative process of assigning volumes to the transportation network and recalculating the skims, which are used in the mode and destination choice model steps.
- **Transportation network.** This includes highway, transit, etc. networks which connect the \acr{TAZ}s to each other.
These networks contain information such as link speed and capacity.
Though the \acr{WFRC} model assigns travel volumes to the network, this research does not compare the model's network assignment results.
However, the network volumes are still used to calculate the travel skims.
- **Model constants and coefficients.** Many model steps, including mode choice, require calibration constants and coefficients.
These are taken directly from the \acr{WFRC} model without modification.

The model is comprised of 5 main steps: household disaggregation, trip generation, trip distribution, mode choice, and network assignment.
Each of these steps has several sub-steps, and there are additional pre- and post-processing steps, but these five steps are the core of the model.
The household disaggregation model uses the land use and socioeconomic data (containing average household sizes, etc.) and estimates the number of households by type, meaning by number of residents and number of workers.
An additional part of this step is a vehicle ownership model, which further disaggregates the households by number of owned vehicles.
The remaining four steps proceed similarly to any other trip-based model, estimating trips from each household type by purpose, time of day, origin/destination, and mode choice, and assigning these trips to the network(s).

The final output(s) of the \acr{WFRC} model include network volumes and performance measures.
This includes loaded the network skims, which can then be used in subsequent iterations of the destination and mode choice models.
Additionally, the model outputs intermediate results after each main model step.
This includes disaggregated household data, trip generation estimates, and origin-destination matrices counting trips by purpose, mode, and time of day.

## ActivitySim

<!-- The open-source ActivitySim [@cite] is used as our representative \acr{ABM}. -->
<!-- ActivitySim was developed.... -->
<!-- Many agencies currently use a travel demand model based on ActivitySim in some form [@cite]. -->
<!-- another sentence. -->

<!-- Overview of asim model (inputs, outputs, etc.) (also basic flowchart, tours/trips/etc) -->

<!-- We are getting the input zonal SE data form the WFRC model and using that as targets for our synthetic population generation (via PopulationSim). -->
<!-- We used PopulationSim [@cite] to create the synthetic population to use in ActivitySim. -->
<!-- PopulationSim requires two main categories of inputs: a seed and controls. -->
<!-- The seed comes from XXXXX, and contains a sample of actual (anonymized) individuals and households in the Wasatch Front region. -->
<!-- PopulationSim assigns copies of these seeds to the area in a way that complies with the controls. -->
<!-- The controls contain targets for the synthetic population in categories such as number of households, **et al**. -->
<!-- For this we used a combination of the \acr{TAZ}-level socioeconomic data present in the \acr{WFRC} model and Census tract-level data. -->

## Initial Model Comparison

In order to reasonably compare the two models, the initial "baseline" scenarios in each should have relatively similar results.
Certainly this applies to the predicted trips by purpose, mode, length, etc., but it is also critical to ensure the synthetic population generated by PopulationSim matches the socioeconomic data used in the WFRC/MAG model.
Though many controls such as household size and income are specified from the Census at the tract level, the WFRC socioeconomic data contains this information as well.
Ideally, then, variables such as number of households per TAZ and TAZ median income should be similar between the two sources.

The data from both the Census and the WFRC/MAG model become geography-specific control totals which PopulationSim tries to match, with different weights assigned to each control.
Number of households, for example, was given at the TAZ level from the WFRC data and weighted very heavily, so there is no difference in number of households in each TAZ between the synthetic population and the WFRC data.
Total population, on the other hand, is a control from the Census data, and is in fact specified at the region geography level, and so there is some discrepancy between the synthetic population and the WFRC data (see @fig-totpop-comp).
However, these discrepancies are generally relatively minor, especially in more populated areas.

```{r}
#| label: fig-totpop-comp
#| fig-cap: Comparison of population by TAZ between WFRC data and PopulationSim synthetic population.

targets::tar_read(pop_comp_map)
```

Another important comparison to make is of household income, as this can significantly affect travel behavior.
Instead of simply comparing average household income by TAZ, it is more useful to compare the number of households in each binned income group.
Both the WFRC and ActivitySim models use income *groups* as a variable in their model steps rather than the actual income.
The income groups for the WFRC model are given in @tbl-wfrc-income-groups.

```{r}
#| label: tbl-wfrc-income-groups
#| tbl-cap: income groups

targets::tar_read(income_groups) %>% 
  kbl(
    col.names = c("Income Group", "Income Range"),
    align = "c",
    booktabs = TRUE,
    linesep = ""
  ) %>% 
  kable_styling()
```

@fig-inc-groups-map shows a comparison of the number of households in each income group between the WFRC household disaggregation model and the synthetic population generated by PopulationSim.
There is a definite trend of the synthetic population to over-predict low-income households and over-predict high-income households compared to the WFRC disaggregation model.
However, especially since this research is focused more on a methodological comparison between the two models rather than a comparison of results, the discrepancies are not too severe to be concerned about.

```{r}
#| label: fig-inc-groups-map
#| fig-cap: income groups

targets::tar_read(inc_groups_map)
```



```{r}
#| label: tbl-mode-split
#| tbl-cap: mode split

targets::tar_read(mode_split_comp) %>% 
  kbl(
    col.names = c("Purpose", "Mode", "ActivitySim predicted trips", "WFRC/MAG predicted trips", "% Error", "% Error (scaled to WFRC/MAG predicted trips)"),
    digits = c(0,0,0,0,1,1),
    align = "c"
  ) %>% 
  kable_styling() %>% 
  column_spec(3, "4cm") %>% 
  column_spec(4, "4cm") %>% 
  column_spec(6, "6cm") %>% 
  collapse_rows(1, valign = "middle", latex_hline = "major")
```

```{r}
#| label: fig-tlfd-comp
#| fig-cap: tlfd comparison

targets::tar_read(tlfd_comp_plot)
```


<!-- Compare TLFD, hopefully similar. -->

<!-- They are similar! -->
<!-- NOTE: The WFRC model multiplies trips by 100 in mode choice because there are a lot of fractional trips that would be lost to rounding otherwise. -->
<!-- We don't need to do that here (R can handle more than 2 decimal places), so we divided the trips by 100 to get the actual amount. -->

```{r}
#| echo: false

# targets::tar_read(combined_trips) %>%
#   group_by(model, mode, purpose) %>%
#   summarise(trips = sum(trips)) %>%
#   pivot_wider(names_from = model, values_from = trips) %>%
#   arrange(purpose, mode) %>%
#   mutate(error = asim/wfrc - 1) %>%
#   relocate(purpose)
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| error: false

# targets::tar_read(combined_trips) %>%
#   ggplot() +
#   geom_density(aes(x = distance, weight = trips, color = model)) +
#   facet_grid(
#     vars(mode), vars(purpose),
#     switch = "y",
#     scales = "free") +
#   xlim(0,25) +
#   labs(y = "Kernel density", x = "Distance (mi)", color = "Model")
```



## policy scenarios

We will model the following scenarios:
- Land use
- Transit
- WFH

## Computational resources?

Describe BEAST and the BYU Supercomputer.
